---
title       : Global Socioeconomic Evaluation
subtitle    : Application in Shiny
author      : Bamini Balaji
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [htmlwidgets]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Overview

This presentation is part of the Developing Data Products Coursera course project submission.

The goal of this project was to create a shiny application to visualize worldwide socioeconomic indicators over time. 

Data was sourced from the [Kaggle dataset](https://www.kaggle.com/sdorius/countryses) 

The shiny application can be found [here](https://bbamini.shinyapps.io/socioeconomic/)

The Github repository containing the code for the shiny app can be found [here](https://github.com/bbamini/Data-Products-Project)

--- .class #id 

## Kaggle Data Set

Three socioeconomic indicators were evaluated:

- GDP per capita
- Socioeconomic status score
- Years of education a typical adult underwent

Here are the years that were considered in this evaluation
```{r, echo=FALSE}
setwd('..')
socio <- read.csv("GLOBCSES.csv", header = TRUE, colClasses = c("integer", "factor", "factor", "numeric", "numeric", "factor", "numeric", "numeric", "factor", "factor"))
socio$year[1:14]
```

Here are the number of countries that were evaluated:
```{r, echo=FALSE}
nlevels(socio$country)
```

---

## Table Visualization

The Best and Worst tab in the app displays the 10 countries with the lowest values and 10 countries with the highest values for the economic indicator in a given year. 

Here is the output for the 5 countries with highest GDP per capita in 2010 
```{r, echo=FALSE, warning=FALSE, comment=FALSE, message=FALSE}
interest <- socio[, c(3, 4, 5, 7, 8)]
yr_interest <- subset(interest, year == 2010)
yr_interest[order(yr_interest$gdppc, decreasing = TRUE),][1:5,c(1,3,4,5)]
```

```{r, echo=FALSE, warning=FALSE, comment=FALSE, message=FALSE}
library(ggplot2)
library(plotly)
library(webshot)
```

---

## Map Visualization

Here is the output for the global distribution of GDP per capita in 2010 

```{r echo=TRUE, warning=FALSE, comment=FALSE, message=FALSE, results=FALSE}
g <- list(showframe = FALSE, showcoastlines = TRUE, projection = list(type = 'Mercator'))
p <- plot_geo(yr_interest) %>% add_trace(z = ~gdppc, color = ~gdppc, text = yr_interest$country, locations = yr_interest$country, locationmode = "country names") %>% colorbar(title = "GDP Per Capita", limits = c(500, 50000)) %>% layout(geo = g)
htmlwidgets::saveWidget(as.widget(p), file = "demo.html")
```
<iframe src="demo.html" style="position:absolute;height:70%;width:70%"></iframe>
